---
title: "BSMM-lab-7"
subtitle: "BSMM 8740 Fall 2023"
author: "Add your name here"
date: "Add the date here"
format: html
editor: visual
self-contained: true
---

::: callout-note
## REMINDER:

Be sure to edit this document (see above) to include your [**name**]{.underline} (and the [**date**]{.underline})

Before you wrap up the assignment, make sure all documents are updated on your GitHub repo (i.e. save, stage, commit and **push**).

Remember -- you do [**not**]{.underline} have to turn in an \*.html file. I will be pulling your work directly from your repository on the [course github site](https://github.com/BSMM-8740-Fall-2023).
:::

## Introduction

In today's lab, you'll practice building `workflows` with `recipes`, `parsnip` models, `rsample` cross validations, and model comparison in the context of timeseries data.

## Packages

```{r}
#| message: false
#| echo: false
library(magrittr)   # the pipe
library(tidyverse)  # for data wrangling + visualization
library(tidymodels) # for modeling
library(modeltime)  # for modeling ts
library(ggplot2)    # for plotting
# set the efault theme for plotting
theme_set(theme_bw(base_size = 18) + theme(legend.position = "top"))
```

## The Data

Today we will be using electricity demand data, based on a paper by James W Taylor:

> Taylor, J.W. (2003) Short-term electricity demand forecasting using double seasonal exponential smoothing. Journal of the Operational Research Society, 54, 799-805.

The data can be found in the `timetk` package as `timetk::taylor_30_min`, a tibble with demensions: 4,032 x 2

-   `date`: A date-time variable in 30-minute increments

-   `value`: Electricity demand in Megawatts

```{r}
#| eval: true
data <- timetk::taylor_30_min
```

## Exercise 1: EDA

Plot the data using the functions `timetk::plot_time_series`, `timetk::plot_acf_diagnostics` (using 100 lags), and `timetk::plot_seasonal_diagnostics`.

::: {.callout-note appearance="simple" icon="false"}
## YOUR ANSWER:

```{r}

```
:::

## Exercise 2: Time scaling

The raw data has 30 minutes intervals between data points. Downscale the data to 60 minute intervals, using `timetk::summarise_by_time`, revising the electricity demand (value) variable by adding the two 30-minute intervals in each 60-minute interval. Assign the downscaled data to the variable `taylor_60_min`.

::: {.callout-note appearance="simple" icon="false"}
## YOUR ANSWER:

```{r}
set.seed(8740)
taylor_60_min <- ?
```
:::

## Exercise 3: Training and test datasets

1.  Split the new (60 min) time series into training and test sets using `timetk::time_series_split`
    1.  set the training period ('initial') to '2 months' and the assessment period to '1 weeks'
2.  Prepare the data resample specification with `timetk::tk_time_series_cv_plan()` and plot it with `timetk::plot_time_series_cv_plan`
3.  Separate the training and test data sets using `rsample` .

::: {.callout-note appearance="simple" icon="false"}
## YOUR ANSWER:

```{r}

```
:::

## Exercise 4: recipes

1.  Create a base recipe (base_rec) using the formula `value ~ date` and the training data. This will be used for non-regression models

2.  Create a recipe (lm_rec) using the formula `value ~ .` and the `training` data. This will be used for regression models. For this recipe:

    1.  add time series signature features using `timetk::step_timeseries_signature` with the appropriate argument,
    2.  add a step to select the columns `value`, `date_index.num`, `date_month.lbl`, `date_wday.lbl`, `date_hour` ,
    3.  add a normalization step targeting `date_index.num` ,
    4.  add a step to mutate `date_hour`, changing it to a factor,
    5.  add a step to one-hot encode nominal predictors.

::: {.callout-note appearance="simple" icon="false"}
## YOUR ANSWER:

```{r}
base_rec <- ?
  
lm_rec <- ?
```
:::

## Exercise 5 models

Now we will create a several models to estimate electricity demand, as follows

1.  Create a model specification for an exponential smoothing model using engine 'ets'
2.  Create a model specification for an arima model using engine 'auto_arima'
3.  Create a model specification for a linear model using engine 'glmnet' and penalty = 0.02, mixture = 0.5

::: {.callout-note appearance="simple" icon="false"}
## YOUR ANSWER:

```{r}
model_ets <- ?
  
model_arima <- ?
  
model_lm <- ?
```
:::

## Exercise 6 model fitting

Create a workflow **for each model** using `workflows::workflow`.

1.  Add a recipe to the workflow
    1.  the linear model uses the `lm_rec` recipe created above
    2.  the `ets` and `arima` models use the `base_rec` recipe created above
2.  Add a model to each workflow
3.  Fit with the training data

::: {.callout-note appearance="simple" icon="false"}
## YOUR ANSWER:

```{r}
workflow_fit_ets <- ?
  
workflow_fit_arima <- ?
  
workflow_fit_lm <- ?
```
:::

::: render-commit-push
This is a good place to render, commit, and push changes to your remote lab repo on GitHub. Click the checkbox next to each file in the Git pane to stage the updates you've made, write an informative commit message, and push. After you push the changes, the Git pane in RStudio should be empty.
:::

## Exercise 7: calibrate

In this exercise we'll use the **testing data** with our fitted models.

1.  Create a table with the fitted workflows using `modeltime::modeltime_table`
2.  Using the table you just created, run a calibration on the **test data** with the function `modeltime::modeltime_calibrate`.
3.  Compare the accuracy of the models using the `modeltime::modeltime_accuracy()` on the results of the calibration

::: {.callout-note appearance="simple" icon="false"}
## YOUR ANSWER:

```{r}
model_tbl <- modeltime::modeltime_table( ? )

calibration_tbl <- ?
  
calibration_tbl %>% ?
```

::: callout-important
Which is the best model by the **rmse** metric?
:::
:::

## Exercise 8: forecast - training data

Use the calibration table with `modeltime::modeltime_forecast` to graphically compare the fits to the training data with the observed values.

::: {.callout-note appearance="simple" icon="false"}
## YOUR ANSWER:

```{r}

```
:::

## Exercise 9: forecast - future

Now refit the models using the **full data set** (using the calibration table and `modeltime::modeltime_refit`). Save the result in the variable **refit_tbl**.

1.  Use the refit data in the variable refit_tbl, along with `modeltime::modeltime_forecast` and argument `h` = '2 weeks' (remember to also set the `actual_data` argument). This will use the models to forecast electricity demand two weeks into the future.
2.  Plot the forecast with `modeltime::plot_modeltime_forecast`

::: {.callout-note appearance="simple" icon="false"}
## YOUR ANSWER:

```{r}
refit_tbl <-  ?
```
:::

## Submission

::: callout-warning
Before you wrap up the assignment, make sure all documents are saved, staged, committed, and pushed to your repository on the course github site.

Remember -- you do **not** have to turn in an \*.html file. I will be pulling your work directly from your repository on the course website.
:::

## Grading

Total points available: 30 points.

| Component | Points |
|-----------|--------|
| Ex 1 - 9  | 30     |
